### 2010-03-29 papers/icfp-2010-desc/paper_type_theory.tex
Desc@ICFP: we use Lisp notation for EnumU.
Desc@ICFP: cite other users of bidirectional type-checking, oriented boilerplate scrapping.

As a motivating example, we are now going to extend the type theory
with a notion of finite set. One could argue that there is no need for
such an extension: finite sets, just as any data-structure, can be
encoded inside the type theory. A well-known example of such encoding
is the Church encoding of natural numbers, which is isomorphic to
finite sets.

% Another example is Martin-Löf's W-types~\cite{}.

\note{Shall we talk about W-types encoding?}

@@

\note{Did I got the coding presentation vs. coding representation
  story right?}

@@

The specification of finite sets is presented in
Figure~\ref{fig:typing-finite-set}. It is composed of three
components. First, we define tags as inhabitants of the $\UId$ type. A
tag is solely an informative token, used for diagnostic
purposes. Finite sets inhabits the $\EnumU$ type. Unfolding the
definition, we get that a finite set is a list of tags. Finally,
elements of a finite set $\V{u}$ belong to the corresponding $\EnumT{\V{u}}$
type. Intuitively, it corresponds to an index -- a number -- pointing
to an element of $\V{u}$.

@@

A second consequence concerns the definition of a finite set. We can
abstract away the details of code and use a familiar syntax
instead. Hence, we use the syntax $[t_1 \ldots t_n]$ to denote the
finite sets composed of the named elements $t_1$, \ldots, $t_n$. The
translation to elements of $\EnumU$ is straightforward.

@@

In this section, we have developed a type propagation system based on
bidirectional type-checking. Using bidirectional type-checking as a
boilerplate scrapper is a well-known technique~\cite{who?} \note{I
  would need some citation of people using bidirectional presentation
  of their type theory}. In our case, we have shown how to instrument
bidirectionality to rationalise the expressivity of dependent
types. We have illustrated our approach with finite sets. We have
abstracted away the low-level presentation of finite sets, offering a
convenient syntax instead.

>>>>

As a motivating example, we are now going to extend the type theory
with a notion of finite set. One could argue that there is no need for
such an extension: finite sets, just as any data-structure, can be
encoded inside the type theory. A well-known example of such encoding
is the Church encoding of natural numbers, which is isomorphic to
finite sets. \note{Shall we talk about W-types encoding?}

@@

\note{Did I got the coding presentation vs. coding representation
  story right? No.}

@@

The specification of finite sets is presented in
Figure~\ref{fig:typing-finite-set}. It is composed of three
components. First, we define tags as inhabitants of the $\UId$ type. A
tag is solely an informative token, used for diagnostic
purposes. Finite sets inhabits the $\EnumU$ type. Unfolding the
definition, we get that a finite set is a list of tags. Finally,
elements of a finite set $\V{u}$ belong to the corresponding $\EnumT{\V{u}}$
type. Intuitively, it corresponds to an index -- a number -- pointing
to an element of $u$.

@@

A second consequence concerns the definition of a finite set. We can
abstract away the details of code and use a familiar syntax
instead. Hence, we use the syntax $[t_1 \ldots t_n]$ to denote the
finite sets composed of the named elements $t_1$, \ldots, $t_n$. This
syntax is reminiscent to Lisp notation for lists. The translation to
elements of $\EnumU$ is straightforward.

@@

In this section, we have developed a type propagation system based on
bidirectional type-checking. Using bidirectional type-checking as a
boilerplate scrapper is a well-known
technique~\cite{pierce:bidirectional-tc,
  xi:bidirectional-tc-bound-array, chlipala:strict-bidirectional-tc}
\note{Everybody ok with the citations?}. In our case, we have shown
how to instrument bidirectionality to rationalise the expressivity of
dependent types. We have illustrated our approach with finite sets. We
have abstracted away the low-level presentation of finite sets,
offering a convenient syntax instead.
